Deep mutational scanning of hemagglutinin helps predict evolutionary fates of human H3N2 influenza variants

### 2018-06-24 paper/manuscript.tex
Split mut validation fig into separate figs, update captions

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.9\textwidth]{figs/supp_diffprefs/diffprefs.pdf}}
\caption{\label{suppfig:diffprefs_logoplot}
{\bf Sites where there are strong differences between our experimental measurements and the amino-acid frequencies among natural HA sequences.}
We calculated the distance between our H3 measurements and the alignment frequencies using the same approach as in Figure~\ref{fig:distance_distribution}, but using the alignment frequencies in place of the H1 preferences.
For each site, the height of each letter above or below the line indicate how much more or less preferred that amino acid is in our experiments compared to its frequency nature.
The overlays show the same information as in Figure~\ref{fig:logoplot} (domain and wildtype amino acid).
The sites are in H3 numbering.
The HA alignment used to calculate the natural frequencies is the same one used in Table~\ref{tab:phydms}.
\comment{Juhye: the notebook that creates this has a bug, as it never creates the \texttt{H3\_alignment\_frequences\_copy.csv}. 
This must be something you did by hand but then didn't include the code for in the notebook: fix this!
Also, it might be nice if you added a bit more explanation to this section of the Jupyter notebook.
}
}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.8\textwidth]{figs/supp_point_mut_validation/point_mut_validation.pdf}}
\caption{
\comment{Just show the first panel (A) here}
{\bf Validation that the Perth/2009 HA is somewhat tolerant to mutation of the canonical start codon.}
We created variants of the Perth/2009 HA in which the canonical start codon was mutated (amino-acid mutation M(-16)K, codon mutation \comment{what was it}), a conserved cysteine was mutated to alanine (C52A, codon mutation \comment{what was it?}), and the same cysteine was synonymously mutated (C52C, codon mutation \comment{what was it}).
We selected these mutations because our deep mutational scanning results in Figure~\ref{fig:logoplot} surprisingly suggests that the start codon (position -16) is fairly tolerant of mutations, but that site 52 is highly intolerant of mutation to anything other than cysteine.
The synonymous C52C mutation is a negative control mutation that is not expected to have any effect.
We then used reverse genetics to generate viruses carrying the wildtype HA or each point mutant, with GFP packaged in the PB1 segment to enable easy titering by flow cytometry~\cite{bloom2010permissive,hooper2013mutant}.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
As expected, the C52A virus is essentially non-viable, while the C52C mutation is roughly equivalent to wildtype.
the M(-16)K mutation is only moderately attenuated, validating that the canonical start codon is not completely essential for viral growth.
}
\label{suppfig:startcodon_validation}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.8\textwidth]{figs/supp_point_mut_validation/point_mut_validation.pdf}}
\caption{\label{suppfig:glycan_validation}
\comment{Just show the third panel (C) here, removing the cysteine mutants and adjusting the y-axis to a more appropriate for just these mutants.}
{\bf Validation of viral variants with mutations at N-linked glycosylation motifs.}
We created variants of the Perth/2009 HA with mutations that eliminated N-linked glycosylation motifs (Asn-Xaa-Ser/Thr) at asparagine residues 22, 38, and 285 (these are mutations T24F, T40V, and S287A, respectively).
We then used reverse genetics to generate viruses carrying the wildtype HA or each of these mutants.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
The viruses with mutations at the motifs at residues 22 and 38 reached titers at least as high as wildtype, whereas the virus with a mutation to the motif at residue 285 was modestly attenuated.
}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.8\textwidth]{figs/supp_point_mut_validation/point_mut_validation.pdf}}
\caption{\label{suppfig:TM_validation}
\comment{Just show TM mutant here, similar to previous figure}
{\bf Validation of the mutational tolerance of a site in the transmembrane domain.}
We created a variant of the Perth/2009 HA with a transmembrane domain mutation, C199(HA2)K, at a site that our deep mutational scanning suggests should be highly mutationally tolerant (Figure~\ref{fig:logoplot}).
We then used reverse genetics to generate viruses carrying the wildtype HA or this mutants
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
The virus the mutation reached titers comparable to wildtype.
}
\end{suppfigure}

>>>>

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.9\textwidth]{figs/supp_diffprefs/diffprefs.pdf}}
\caption{\label{suppfig:diffprefs_logoplot}
{\bf Sites where there are strong differences between our experimental measurements and the amino-acid frequencies among natural HA sequences.}
We calculated the distance between our H3 measurements and the alignment frequencies using the same approach as in Figure~\ref{fig:distance_distribution}, but using the alignment frequencies in place of the H1 preferences.
For each site, the height of each letter above or below the line indicate how much more or less preferred that amino acid is in our experiments compared to its frequency nature.
The overlays show the same information as in Figure~\ref{fig:logoplot} (domain and wildtype amino acid).
The sites are in H3 numbering.
The HA alignment used to calculate the natural frequencies is the same one used in Table~\ref{tab:phydms}.
}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.4\textwidth]{figs/supp_point_mut_validation/startcodon_titers.pdf}}
\caption{
{\bf Validation that the Perth/2009 HA is somewhat tolerant to mutation of the canonical start codon.}
We created variants of the Perth/2009 HA in which the canonical start codon was mutated (amino-acid mutation M(-16)K, codon mutation ATG(-16)AAA), a conserved cysteine was mutated to alanine (C52A, codon mutation TGC52GCA), and the same cysteine was synonymously mutated (C52C, codon mutation TGC52TGT).
We selected these mutations because our deep mutational scanning results in Figure~\ref{fig:logoplot} surprisingly suggests that the start codon (position -16) is fairly tolerant of mutations, but that site 52 is highly intolerant of mutation to anything other than cysteine.
The synonymous C52C mutation is a negative control mutation that is not expected to have any effect.
We then used reverse genetics to generate viruses carrying the wildtype HA or each point mutant, with GFP packaged in the PB1 segment to enable easy titering by flow cytometry~\cite{bloom2010permissive,hooper2013mutant}.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
As expected, the C52A virus is essentially non-viable, while the C52C mutation is roughly equivalent to wildtype.
the M(-16)K mutation is only moderately attenuated, validating that the canonical start codon is not completely essential for viral growth.
}
\label{suppfig:startcodon_validation}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.4\textwidth]{figs/supp_point_mut_validation/glycan_titers.pdf}}
\caption{\label{suppfig:glycan_validation}
{\bf Validation of viral variants with mutations at N-linked glycosylation motifs.}
We created variants of the Perth/2009 HA with mutations that eliminated N-linked glycosylation motifs (Asn-Xaa-Ser/Thr) at asparagine residues 22, 38, and 285 (these are mutations T24F, T40V, and S287A, respectively).
We then used reverse genetics to generate viruses carrying the wildtype HA or each of these mutants.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
The viruses with mutations at the motifs at residues 22 and 38 reached titers at least as high as wildtype, whereas the virus with a mutation to the motif at residue 285 was modestly attenuated.
}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.45\textwidth]{figs/supp_point_mut_validation/tm_titers.pdf}}
\caption{\label{suppfig:TM_validation}
{\bf Validation of the mutational tolerance of a site in the transmembrane domain.}
We created a variant of the Perth/2009 HA with a transmembrane domain mutation, C199(HA2)K, at a site that our deep mutational scanning suggests should be highly mutationally tolerant (Figure~\ref{fig:logoplot}).
We then used reverse genetics to generate viruses carrying the wildtype HA or this mutants
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
The virus the mutation reached titers comparable to wildtype.
}
\end{suppfigure}
